---
title: "Vectors"
output: html_notebook
---

# Plotting vectors

A vector:

```{r}
vector <- c(1, 2)
transposedVector <- t(vector)

plot(NA, xlim=c(0,5), ylim=c(0,5), xlab="X", ylab="Y", lwd=2)
arrows(0, 0, vector[1], vector[2], length = 0.35, col=3,lwd=2)
```

```{r}
library(ggplot2)
df <- data.frame(x=c(0),y=c(0), dx=vector[1], dy=vector[2])
ggplot(data=df, aes(x=x, y=y)) + geom_segment(aes(xend=x+dx, yend=y+dy), arrow = arrow(length = unit(0.3,"cm")))
```

# Multiplying a vector by a number

```{r}
originalVector <- c(2, 3)
multiplied <- 2 * originalVector

plot(NA, xlim=c(0,10), ylim=c(0,10), xlab="X", ylab="Y", lwd=2)
arrows(0, 0, originalVector[1], originalVector[2], length = 0.35, col=2,lwd=2)
text(originalVector[1]+1.1, originalVector[2], labels = c("originalVector"), col=2)
arrows(0, 0, multiplied[1], multiplied[2], length = 0.35, col=3,lwd=2)
text(multiplied[1]+1.1, multiplied[2], labels = c("multiplied"), col=3)
```

# Sum of vectors

```{r}
a <- c(1, 4)
b <- c(3, 1)
c <- a + b

plot(NA, xlim=c(0,10), ylim=c(0,10), xlab="X", ylab="Y", lwd=2)
arrows(0, 0, a[1], a[2], length = 0.35, col=2,lwd=2)
text(a[1]+.2, a[2], labels = c("a"), col=2)
arrows(0, 0, b[1], b[2], length = 0.35, col=3,lwd=2)
text(b[1]+.1, b[2], labels = c("b"), col=3)
arrows(0, 0, c[1], c[2], length = 0.35, col=1,lwd=2)
text(c[1]+.8, c[2], labels = c("c = a + b"), col=1)
```

Alternatively, we can start drawing the vector _b_ at the end of vector _a_:

```{r}

plot(NA, xlim=c(0,10), ylim=c(0,10), xlab="X", ylab="Y", lwd=2)
arrows(0, 0, a[1], a[2], length = 0.35, col=2,lwd=2)
text(a[1]+.2, a[2], labels = c("a"), col=2)
arrows(a[1], a[2], a[1] + b[1], a[2] + b[2], length = 0.35, col=3,lwd=2)
text(a[1] + b[1] +.1, a[2] + b[2], labels = c("b"), col=3)
arrows(0, 0, c[1], c[2], length = 0.35, col=1,lwd=2)
text(c[1]+.4, c[2] - 1, labels = c("c = a + b"), col=1)

```

In one way or another, we get the same result _c = (4, 5)_.

# Inner product

The product is a number – not a vector.

```{r}
sum(a * b)
```

# The length (norm) of a vector

```{r}
sqrt(sum(a * a))
```

# The Cross Product

```{r}
a <- c(1, 0, 0)
b <- c(0, 1, 0)

mm <- cbind(a, b)
xy<-sapply(1:3,function(x)det(mm[-x,])*(2*(x%%2)-1))  
xy
```

# References

* [Plotting vectors in a coordinate system with R or python](https://stackoverflow.com/questions/10882336/plotting-vectors-in-a-coordinate-system-with-r-or-python?utm_medium=organic&utm_source=google_rich_qa&utm_campaign=google_rich_qa)
* [vector field visualisation R](https://stackoverflow.com/questions/14936504/vector-field-visualisation-r) (ggplot2)
* [Cross product](https://en.wikipedia.org/wiki/Cross_product) (wiki)
* [R - Compute Cross Product of Vectors](https://stackoverflow.com/questions/36798301/r-compute-cross-product-of-vectors-physics?utm_medium=organic&utm_source=google_rich_qa&utm_campaign=google_rich_qa)
* [vectors cross-product V1 x V2](http://r.789695.n4.nabble.com/vectors-cross-product-V1-x-V2-td3653816.html)
