---
title: "Matrices"
output: html_notebook
---

# Creating a matrix

To create a 3×2 matrix that contains in the first column 2, 3, 4 and the second column 5, 6, 7:

```{r}
A <- matrix(seq(2,7), 3, 2)
A
A[3,] # the third row
A[,2] # the second column
A[3, 2] # the second column of the third row
A[-2,] # all rows but the second one
A[A>2 & A < 7] # all elements of A greater than 2 and smaller than 7
```

To enter a matrix by the rows instead of the columns, then add the argument _byrow=TRUE_ to the command parameters:

```{r}
A <-  matrix(seq(2,7), 3, 2, byrow = TRUE)
A
```

# Combine rows and columns

```{r}
u <- 1:6
v <- 2:7
cbind(u, v)
```

```{r}
rbind(u, v)
```
# Dimensions of a matrix

```{r}
A <- matrix(seq(1:6), 2, 3)
A
dim(A) # dimension of a matrix
length(A) # the number of elements in a matrix
nrow(A) #  the number of rows in a matrix
ncol(A) # the number of columns in a matrix
dimnames(A) # the name of the the dimensions
```

# Matrix operations

```{r}
A <- matrix(c(1:5, 11:14), 3, 3)
B <- matrix(1:9, 3, 3)
A
B
```

## Addition

```{r}
A + B
```

## Subtraction

```{r}
A - B
```

## Multiplication

```{r}
A %*% B
```

## Element-wise multiplication

```{r}
A * B
```

## Transposition

```{r}
t(A)
```

## Diag

To extract the diagonal of a matrix or create a diagonal matrix from a vector, use the command _diag_:

```{r}
diag(A)
```

```{r}
diag(c(1, 2, 3))
diag(diag(A))
```

## Solve

To find the inverse of a matrix or solve a system of equations, use the command _solve_: 

```{r}
solve(A)
```

```{r}
b <- c(2, 3,4)
solve(A, b)
```

## Generalized inverse matrix

Not all square matrices have an inverse.  However all square matrices have an infinite number of generalized inverses. 
A generalized inverse of a square matrix _A_ is a matrix _A-_ satisfying that _A * A- * A = A_. For many practical problems it suffice to find a 
generalized inverse.

```{r}
library(MASS)
ginv(A)
```

## Crossprod

To create the crossproduct t(A)A in an efficient manner, use the command _crossprod_:

```{r}
crossprod(A)
t(A) %*% A
```

## Sum

```{r}
sum(A)
```

## Prod

To multiply the elements of a vector or a matrix, use the command _prod_:

```{r}
prod(A)
prod(1:5)
```

## Cumsum

To create a vector of cumulative sums of the elements of a vector, use the command _cumsum_. If used on a matrix, the matrix is first changed to a vector.

```{r}
cumsum(A)
cumsum(1:5)
```

## Cumprod

To create a vector of cumulative products of the elements of a vector, use the command _cumprod_:

```{r}
cumprod(A)
cumprod(1:5)
```


## Min and Max

```{r}
min(A)
max(A)
```

## Det

To calculate the determinant of a square matrix, use the command _det_:

```{r}
det(A)
```

# References

* [Examples of Using R with Linear Algebra](http://jekyll.math.byuh.edu/courses/m343/handouts/rlinalg.pdf)
* [Linear algebra in R](https://www.math.uh.edu/~jmorgan/Math6397/day13/LinearAlgebraR-Handout.pdf)
* [Matrices and Linear Algebra in R](http://faculty.business.utsa.edu/vdeolive/STA6023/matricesR.pdf)
* [Introductory linear algebra with R](http://bendixcarstensen.com/APC/linalg-notes-BxC.pdf)
* [R and Linear Algebra](https://www.r-bloggers.com/r-and-linear-algebra/)
* [Principal Component Analysis 4 Dummies: Eigenvectors, Eigenvalues and Dimension Reduction](https://georgemdallas.wordpress.com/2013/10/30/principal-component-analysis-4-dummies-eigenvectors-eigenvalues-and-dimension-reduction/)
* [What are eigenvectors and eigenvalues?](http://www.visiondummy.com/2014/03/eigenvalues-eigenvectors/)
* [A geometric interpretation of the covariance matrix](http://www.visiondummy.com/2014/04/geometric-interpretation-covariance-matrix/)
* [Introduction to Principal Component Analysis](https://tgmstat.wordpress.com/2013/11/21/introduction-to-principal-component-analysis-pca/)
* [Computing and visualizing PCA in R](https://www.r-bloggers.com/computing-and-visualizing-pca-in-r/)
* [Principal Component Analysis in R](https://www.r-bloggers.com/principal-component-analysis-in-r/)
* [5 functions to do Principal Components Analysis in R](http://www.gastonsanchez.com/visually-enforced/how-to/2012/06/17/PCA-in-R/)
* [Practical Guide to Principal Component Analysis in R & Python](https://www.analyticsvidhya.com/blog/2016/03/practical-guide-principal-component-analysis-python/)
* [Singular-value decomposition](https://en.wikipedia.org/wiki/Singular-value_decomposition)

