---
title: "Ggplot2 Overview"
output: 
  html_notebook:
    toc: true
    toc_depth: 3
---

# Getting started

## Getting started with ggplot2

In this chapter we’ll mostly use one data set that’s bundled with ggplot2:
mpg . It includes information about the fuel economy of popular car models in 1999 and 2008, collected by the US Environment Protection Agency.

You can access the data by loading ggplot2

```{r}
library(ggplot2)
mpg
```

Every ggplot2 plot has three key components:

1. __data__,
2. A set of __aesthetic mappings__ between variables in the data and visual properties, and
3. At least one layer which describes how to render each observation. Layers
are usually created with a __geom function__.

Here’s a simple example:

```{r}
ggplot(mpg, aes(x = displ, y = hwy)) +
    geom_point()
```

Almost every plot maps a variable to x and y , so naming these aesthetics
is tedious, and first two unnamed arguments to aes() will be mapped to x
and y . That means that following code is identical to the example above:

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point()
```

```{r}
ggplot(economics, aes(date, unemploy)) + geom_line()
```

```{r}
ggplot(mpg, aes(cty)) + geom_histogram()
```

### Colour, size, shape and other aesthetic attributes

```{r}
ggplot(mpg, aes(displ, cty, colour = class)) +
  geom_point()
```

If you want to set an aesthetic to a fixed value, without scaling it, do so
in the individual layer outside of aes() . Compare the following two plots:

```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point(aes(colour = "blue"))
ggplot(mpg, aes(displ, hwy)) + geom_point(colour = "blue")
```

### Faceting

```{r}
ggplot(mpg, aes(displ, hwy)) +
geom_point() +
facet_wrap(~class)
```

### Plot geoms

You might guess that by substituting geom_point() for a different geom func-
tion, you’d get a different type of plot:

* __geom_smooth()__ fits a smoother to the data and displays the smooth and its standard error.
* __geom_boxplot()__ produces a box-and-whisker plot to summarise the distri-
bution of a set of points.
* __geom_histogram()__ and __geom_freqpoly()__ show the distribution of continuous
variables.
* __geom_bar()__ shows the distribution of categorical variables.
* __geom_path()__ and geom_line() draw lines between the data points. A line
plot is constrained to produce lines that travel from left to right, while
paths can go in any direction.

```{r}
ggplot(mpg, aes(displ, hwy)) +
geom_point() +
geom_smooth()
```

```{r}
ggplot(mpg, aes(drv, hwy)) + geom_boxplot()
ggplot(mpg, aes(drv, hwy)) + geom_violin()
```

```{r}
ggplot(mpg, aes(hwy)) +
geom_freqpoly(binwidth = 2.5)
ggplot(mpg, aes(hwy)) +
geom_histogram(binwidth = 2.5)
```

```{r}
ggplot(mpg, aes(displ, colour = drv)) +
geom_freqpoly(binwidth = 0.5)
```

### Time series with line and path plots

```{r}
year <- function(x) as.POSIXlt(x)$year + 1900
ggplot(economics, aes(unemploy / pop, uempmed)) +
geom_path(colour = "grey50") +
geom_point(aes(colour = year(date)))
```

### Modifying the axes

```{r}
ggplot(mpg, aes(cty, hwy)) +
geom_point(alpha = 1 / 3) +
xlab("city driving (mpg)") +
ylab("highway driving (mpg)")
# Remove the axis labels with NULL
ggplot(mpg, aes(cty, hwy)) +
geom_point(alpha = 1 / 3) +
xlab(NULL) +
ylab(NULL)
```

## Toolbox


