---
title: "Exponential Distribution"
output: html_notebook
---

# Introduction

This entire paragraph was taken over from the Introduction to Statistical Thought (written by Michael Lavine; published on June 24, 2010).

It is often necessary to model a continuous random variable X whose density decreases away from 0. Some examples are

* _Customer service_ time on hold at a help line
* _Neurobiology_ time until the next neuron fires
* _Seismology_ time until the next earthquake
* _Medicine_ remaining years of life for a cancer patient
* _Ecology_ dispersal distance of a seed

In these examples it is expected that most calls, times or distances will be short and a few will be long. So the density should be large near _x = 0_ and decreasing as _x_ increases.

```{r}
x <- seq ( 0, 2, length=40 ) # 40 values from 0 to 2
lam <- c ( 2, 1, .2, .1 ) # 4 different values of lambda
y <- matrix ( NA, 40, 4 ) # y values for plotting
for ( i in 1:4 )
  y[,i] <- dexp ( x, 1/lam[i] ) # exponential pdf
matplot ( x, y, type="l", xlab="x", ylab="p(x)", col=1 )
legend ( 1.2, 10, paste ( "lambda =", lam ), lty=1:4, cex=.75 )
```

# Generating Exponential Observations

The function which generates exponential observation in R is called _rexp_.

```{r}
draws <- rexp(400, .1)
mean(draws)
length(draws)
```
# Probability Density

There is a special function in R which calculates the probability density - it is called _dexp_:

```{r}
dexp(45, .1)
```

# Cumulative Density

There is a special function in R which calculates the cumulative density - it is called _pexp_:

```{r}
pexp(45, .1)
```

# Quantiles

Quantile works as an inverse of the cumulative density function, as illustrated below:

```{r}
qexp(pexp(45, .1), .1)
```

Visit [Basic Probability Distributions in R](https://rpubs.com/ggraham412/100906) for more information.

# Quantile-Quantile Plots

As described in the [Q-Q Plot Tutorial](https://intellinexus.wordpress.com/2010/11/29/creating-a-q-q-plot/), The quantile-quantile (q-q) plot is a graphical technique for determining if two data sets come from populations with a common distribution.

The following code compares observations stored in _x.sample_ with those in _y.sample_:

```{r}
library(tidyverse)
x.sample <- rexp(400, .1)
y.sample <- rexp(400, .1)
qqplot(x.sample, y.sample)
abline(0,1)
```

The same test using _ggplot2_:

```{r}
nq <- 31
p <- (1 : nq) / nq - 0.5 / nq
ggplot() + geom_point(aes(x = quantile(x.sample, p), y = quantile(y.sample, p))) +
  geom_abline(intercept = 0, slope = 1)
```

Do the following to compare the _x.sample_ with a Exponential distribution Exp(0.1):

```{r}
xdf <-data.frame(x = x.sample)
ggplot(xdf, aes(sample = x)) + 
  stat_qq(distribution = stats::qexp, dparams = list(rate=.1)) + 
  geom_abline(intercept = 0, slope = 1)
```

# Further Reading

* [A quantile-quantile plot](http://ggplot2.tidyverse.org/reference/geom_qq.html)
* [QQ Plots and PP Plots](http://homepage.divms.uiowa.edu/~luke/classes/STAT4580/qqpp.html)
* [Q-Q Plot Tutorial](https://intellinexus.wordpress.com/2010/11/29/creating-a-q-q-plot/)